---
author: "Pooja Parthasarathy"
title: "class06"
format: pdf
---


## First Function!

Let's write a function that adds some number 

```{r}
add <- function(x,y){
  ##x + 1 ##this is a hardcode 
  ##lets modify to avoid a hardcode:
  x + y
}
```

```{r}
add(c(100,200,300), c(10,20,30)) ## you need not necessarily name everything 
```

```{r}
##for example log will have a DEFAULT VALUE - We can do this for our add function
log(10, base=10)
```

```{r}
add <- function(x,y=1){
  x + y
}
```

```{r}
add(100)
```

>**N.B** Input arguments can be either required or optional, optional ones will contain a default and will be spelled out with an equals

```
name <- function(arg){
  body
}
```

The`sample()` function in R...
```{r}
sample(1:10, size=4)
```

> Q. Return 12 nubers picked ranomldy from the input 1-10 

```{r}
sample(1:10, size=12, replace = TRUE)
```

> Q. Write the code to generate a 12 nuckeotide long DNA Sequence 

```{r}
sample(c('A','T','G','C'), size=12, replace=TRUE)

```

> Generate a first version function called `generate_dna()` that egnerates a user specific length `n` sequence 

```{r}
generate_dna <- function(n=6){
  nucleotides <- sample(c('A','T','G','C'), size=n, replace=TRUE)
  seq_string <- paste0(nucleotide, collapse = "")

  fasta_format <- paste0(">",seq_string)
  return(fasta_format)
   
}


```

> Give the user an option to return FASTA format output sequence or standard multi element vector format 

```{r}
generate_dna <- function(n=6, fasta=TRUE){
  seq_string <- sample(c('A','T','G','C'), size=n, replace=TRUE)
  if(fasta){
      seq_string <- paste0(seq_string, collapse = "")
      cat("Hello....")
      ##fasta_format <- paste0(">",seq_string)
  }
  else{
    cat("is it me you are looking for...")
  }
  return(seq_string)
   
}

generate_dna(n=88, fasta = T)

```

> Generate a function called `generate_protein()`

```{r}

generate_protein <- function(n=6, fasta=TRUE){
  aa20 <- c(
  "A", "R", "N", "D", "C",
  "Q", "E", "G", "H", "I",
  "L", "K", "M", "F", "P",
  "S", "T", "W", "Y", "V"
)
  
 seq_string = sample(aa20, size=n, replace=TRUE)
  
  if(fasta){
      seq_string <- paste0(seq_string, collapse = "")
     ## cat("Hello....")
  }
  else{
    ##cat("is it me you are looking for...")
  }
 
  return(seq_string)
}

generate_protein(12)
```
```{r}
for(i in 6:12){
  cat(">",i,"\n",  sep = "")
  cat(generate_protein(i), "\n", sep = "")
}
```

> trying it all in one for fun

```{r}

generate_protein <- function(n=6, fasta=TRUE){
  aa20 <- c(
  "A", "R", "N", "D", "C",
  "Q", "E", "G", "H", "I",
  "L", "K", "M", "F", "P",
  "S", "T", "W", "Y", "V"
)
  

  for(i in 6:12){
   seq_string = sample(aa20, size=i, replace=TRUE)
  
 
  if(fasta){
      seq_string <- paste0(seq_string, collapse = "")
     ## cat("Hello....")
  }
  else{
    ##cat("is it me you are looking for...")
  }
   print(seq_string)
   ++i
  
  }
}

generate_protein(11)
```


