---
title: "Homework Supplement - R & PDB Function"
author: "Pooja Parthasarathy - A17817456" 
format: pdf
---

# Section 1

```{r}
library(bio3d)

s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plot(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plot(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plot(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")


hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)
```

## Q1. What type of object is returned from the read.pdb() function? 
The type is of a list with roughly 8 elements - within which is a large dataframe, & other lists and data types like double -  for s1,s2,s3. It showcases various properties of the protein in terms of its stuctural identity 

## Q2.  What does the trim.pdb() function do?
Out of the large dataframe initially presented and the initial 8 element list, trim as the name suggests, "trims" in terms of the specified chain type and atom elety property/structural state.

## Q3. What input parameter would turn off the marginal black and grey rectangles in theplots and what do they represent in this case?
It is caused by the plot type chosen, plotb3, once we remove that we are able to eliminate any bars that show up due to the residue presence

##Q4. What would be a better plot to compare across the different proteins? 
A multiline plot that draws this same linegraph however has all three proteins displayed across the same index scale in one graph would make comparison easier

##Q5. Which proteins are more similar to each other in their B-factor trends. How could you quantify this?
From the cluster we can observe that based on "similarity" s2 and s3 show a closer lineage, splitting away from s1 whilst s1 shows more variance in the b-factor trends 

## Q6. How would you generalize the original code above to work with any set of input protein structures?


```{r}

library(bio3d)

bfactor_function <- function(pdb_ids, chain="A", elety="CA") {

##to help eliminate repetitive function calls, we can use lapply
  pdbs <- lapply(pdb_ids, read.pdb)
  chains <- lapply(pdbs, function(x) trim.pdb(x, chain=chain, elety=elety))
  bvals <- lapply(chains, function(x) x$atom$b)

  for(i in seq_along(bvals)) {
  ## for each protein in the given list we iterate to plot for the chain, line plot and show the bfactor
    plot(bvals[[i]], sse=chains[[i]], type="l", ylab="Bfactor") 
  }

  #pull it all together into one dataframe
  bmat <- do.call(rbind, bvals)
  hc <- hclust(dist(bmat))
  plot(hc)

  bmat ##return
}
```

```{r}
bfactor_output <- bfactor_function(c("4AKE","1AKE","1E4Y"))
bfactor_output
```

##Questions & Answers for 6 - Function 
- INPUTS-> this function takes in protein database ids in the form of a vector, and return their bfactor plots across indexes
- To use this, you should first give an input your pdb ids as a vector of strings, the function will then:
1) read the ideas using read.pdb across each item in your vector
2) follows this process to further read chains and bvalues (here they are defined as defaults, but you can specify!)
3) It will use a for loop to traverse through each of those values saved from the lapply using indexes and generate a plot 
4) hclust will generate a phylogeny tree to see differences and similarities

