---
title: "class12 - RNA Seq & Databases"
author: "Pooja Parthasarathy | PID: A17817456"
format: pdf
---

## Question 5 
```{r}

mxl <- read.csv("/Users/poojaparthasarathy/Documents/WINTER 2026 UCSD/BIMM143/r wrok/class12/class12/373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
```

```{r}
length(grep("^G\\|G$", mxl$Genotype..forward.strand.))
```

```{r}
table(mxl$Genotype..forward.strand.)
```

```{r}
table(mxl$Genotype..forward.strand.)/nrow(mxl) * 100 

```


> Now we do this analysis but for the GBR population 

```{r}
gbr <- read.csv("/Users/poojaparthasarathy/Documents/WINTER 2026 UCSD/BIMM143/r wrok/class12/class12/373522-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
```

```{r}
table(gbr$Genotype..forward.strand.)

```

```{r}
table(gbr$Genotype..forward.strand.)/nrow(gbr) * 100 

```

This variat is associated with childhoos asthma and is more frequent in the GBR population in comparison to MXL 



## Population scale analysis

One sample is obviously not enough to know what is happening in a population. You are
interested in assessing genetic differences on a population scale. So, you processed about
~230 samples and did the normalization on a genome level. Now, you want to find whether
there is any association of the 4 asthma-associated SNPs (rs8067378…) on ORMDL3
expression

How many samples do we have
```{r}

expr <-  read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```

Q13. From the file read in, we can use the number of rows to determine the number of samples. This results in 462
```{r}
nrow(expr)
```

```{r}
table(expr$geno)
```

```{r}
library(ggplot2)

ggplot(expr) + aes(geno, exp, fill = geno) + 
  geom_boxplot(notch = TRUE)

```
Q14.
Yes, we can observe that there are differences in expression that occur due to different SNPS. Based on the graphs, a higher expression can be noted from A/A SNPs whilst the lowest expression ranges are associated with G/G, additionally bolstered by notch placement which shows obvious difference in the median expression across all three genotypes.
