---
title: "Class 5: Data Viz with ggplot"
author: "Pooja Parthasarathy (PID:A17817456)"
format: pdf
toc: true
---


## Background 

There are lots of ways to make visualization and plots in R. These include, so-called "base R" (like the function `plot()`) and **ggplot2** 

Let's make the same plot with these two graphics systems. We can use inbuilt `cars` dataset.
```{r}
head(cars)
```
```{r}
plot(cars)
```
first you need to install ggplot using the command `install.packages("ggplot2")`

> **N.B** We never run an install of packages in a code chunk otherwise we will reinstall needlesely every time we render our document 

Everytime you want to use an add on packaged we need to load it up with a call to `library()`

```{r}
library(ggplot2)
ggplot(cars)
```
Every ggplot needs at least 3 things: 

1. The **Data** i.e. struff to plot as a dataframe
2. **Aesthetics** that map the data to the plot 
3. **geom_** or geometry of the plot 

```{r}
ggplot(cars) +
  aes(x=speed, y=dist) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  labs(x="Speed (MPH)", y= "Distance(ft)", title = "stopping distance of old cars") + 
  
  theme_dark()
```


## Gene Expression Plot

Read some fata on the effects of GLP-1 inhibitor drug on gene expression values 
```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```
Version 1 plot - start simple by getting some ink on the page!

```{r}
table(genes$State)

ggplot(genes)+
  aes(Condition1,Condition2, col= State)+
  geom_point() + 
  scale_color_manual(values = c("purple","gray","orange")) +
  labs(x="No Drug", y= "Drug Treatment", title = "Gene Expression Response to GLP-1 drug") +
  theme_classic()
```

## Going further with gapminder 

Here we explore the famous `gapminder` dataset with some custom plots 
```{r}
# File location online
url <- "https://raw.githubusercontent.com/jennybc/gapminder/master/inst/extdata/gapminder.tsv"

gapminder <- read.delim(url)
head(gapminder)
```

> Q. how many rows does this datasethave 

```{r}
nrow(gapminder)
```

> Q. how many different countries are in this dataset 

```{r}
table(gapminder$continent)
```



Version 1 plot GDP Vs Life Expectancy for all rows 
```{r}
ggplot(gapminder) +
  aes(x=gdpPercap, y=lifeExp, col=continent) + 
  geom_point() 
```

We want to see plots for each of the continents in ggplot lingo this is called "faceting" 
```{r}
ggplot(gapminder) +
  aes(x=gdpPercap, y=lifeExp, col=continent) + 
  geom_point() +
  facet_wrap(~continent)
```

## First look at the dplyr package

Another add on package with a function called `filter` that we want to use. 
```{r}
library(dplyr)

input <- filter(gapminder, year == 2007 | year == 1977 )

ggplot(input)+
  aes(x=gdpPercap, y=lifeExp, col=continent) + 
  geom_point() +
  facet_wrap(~year)

```

