---
title: "Class07 - Machine Learning I"
author: "Pooja Parthasarathy"
format: pdf
---

#Background

Today we begin out exploratino of some important machine learning method namely **clustering** & **dimensionality reduction**

Let's make up some input data for clustering where we know what the natural "clusters" are. The function `rnorm()` can be useful

```{r}
hist(rnorm(5000, mean = 10))
```

```{r}
tmp <- c(rnorm(30, mean= 3),
         rnorm(30, mean= -3))

x <- cbind(x=tmp, y = rev(tmp))
plot(x)
```

## K-means clustering

The main function in "base R" for k-means clustering is `kmeans()`:

```{r}
km <-  kmeans(x, centers=2)
km
#centers determines the number of clusters you want, the initial requirements are 2 required arguments of x and centers
```

> Q. What component of the results object details the cluster size

```{r}
km$size
```

> Q. What component of the results object details the cluster centers

```{r}
mem <- km$centers
```

> Q. What component of the results object details the cluster membership vector

```{r}
km$cluster
```

> Q. Plot our clustering results with points colored by cluster and also add cluster centers

```{r}
plot(x, col = km$cluster)
points(km$centers, col = "green", pch=15)
```

> Q. run kmeans() again and this time produce 4 clusters and call your result object `k4`

```{r}
k4 <- kmeans(x, centers = 4)
k4

plot(x, col = k4$cluster)
points(k4$centers, col = "purple", pch = 20)
```

The metric

```{r}
km$tot.withinss
k4$tot.withinss
```

```{r}
ans <- NULL

for (i in 1:30){
  ans <- c(ans, kmeans(x, centers = i)$tot.withinss)
}

ans

```

```{r}
plot(ans, typ = "o")
## a scree plot the point of change here the scree point, the point off the cliff minimizes the difference of sum squared
##so here out preferred answer is 2 
```

## Hierarchical clustering

`hclust()` is the main function for hierarchical clustering - a distance matrix/dissimilarity structure is required. You cannot just pass the dataset, so to do this you will need to pass it through `dist()` which generates a distance matrix

```{r}
d <- dist(x)
hc <- hclust(d)
plot(hc)
```

To extract our cluster membership bector from a `hclust()` result object we have to "cut" our tree at a given height to yield separate groups or branches

```{r}
plot(hc) 
abline(h=8, col = "red", lty = 2)
```

To do this we use the `cutree()` function on our `hclust()` object

# PCA of UK Food Data - Hands on lab

Import the dataset of food consumption in the UK

```{r}
x <- read.csv("https://tinyurl.com/UK-foods")
x
```

> Q1. How many rows and columns are in your new data frame named x? What R functions could you use to answer this questions?

```{r}
dim(x)
```

```{r}
rownames(x) <-  x[,1]
```


> Q2. Which approach to solving the ‘row-names problem’ mentioned above do you prefer and why? Is one approach more robust than another under certain circumstances?

To remove the first column I can use the minus index trick, but a better way to do this to avoid overwriting would be to mention it in the initial read csv itself 
```{r}
x <- read.csv("https://tinyurl.com/UK-foods", row.names = 1)
x
```

### Spotting major differences and trends 

It is difficult even in the tiny 17 dimension dataset 

```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```
###Pairs plots 
```{r}
pairs(x, col=rainbow(nrow(x)), pch=16)

```

```{r}
library(pheatmap)

pheatmap( as.matrix(x) )
```

## PCA to the rescue!

The main PCA function in "base R" is `prcomp()`. This function wants the transpose of our food data input (i.e, food as columns and the countries as rows)

```{r}
pca <- prcomp(t(x)) ##this is a transpose
```

```{r}
## measure the variance for each of these new axis (principal) components 
summary(pca)
```
```{r}
attributes(pca)
```
to make one of our main PCA result figures, we turn to `pca$x` the scores along our new PCs. This is called "PC Plot" or "score plot" or "Ordination plot"

```{r}
my_cols <-  c("orange", "red","blue","darkgreen")
```

```{r}
library (ggplot2)
ggplot(pca$x)+
  aes(PC1, PC2) +
  geom_point(col=my_cols)
```

The second major result figrue is called "loadings plot" or "variable contributions plot" or "weight plot" 
```{r}
ggplot(pca$rotation) +
  aes(PC1, rownames(pca$rotation)) +
  geom_col(fill = "steelblue") +
  
    
xlab("PC1 Loading Score") +
  ylab("") +
  theme_bw() +
  theme(axis.text.y = element_text(size = 9))
```
```{r}
```

