---
title: "PCA Class 8 - Machine Learning II"
format: pdf
---

## Quarto



```{r}
wisconsin <-  read.csv("/Users/poojaparthasarathy/Downloads/WisconsinCancer.csv", row.names = 1)
head(wisconsin)
```

> How many obeservations are in this data set?

```{r}
nrow(wisconsin)
```
> How many obeservations are in this data set?

```{r}
length(grep("M", wisconsin$diagnosis))

##sum(wisconsin$diagnosis == "M")
##sum(table(wisconsin$diagnosis))
```
> how many variables/features in the data are suffixed with mean?

```{r}
names <- c(colnames(wisconsin))

length(grep("_mean", names))

```

We need to remove the diagnosis column before we doany further analysis as to not bias the way we come to an answer. It should not impact the way a PCA works
Save it into a vector so htat you can look back at it later on for verifying your results 

```{r}
wisconsin.data <- wisconsin[,-1]
diagnosis <-  wisconsin$diagnosis

```


```{r}
wisc.pr <-  prcomp(wisconsin.data, scale = TRUE)
```

```{r}
attributes(wisc.pr)
```

```{r}
library(ggplot2)

#wisc.pr$x

ggplot(wisc.pr$x) + 
  aes(PC1,PC2, col = diagnosis)+
  geom_point()
```

## Plots 
OUR ABOVE GRAPH SHOWS US THE SCORE PLOT, it can show you multiple different variables but represents key characteristics across various dimensions 

Using varaiance you can observe the variance covered by various PCs
```{r}
pr.var <- wisc.pr$sdev^2
head(pr.var)
```


```{r}
library(factoextra)
fviz_eig(wisc.pr, addlabels = TRUE)
```



#4 Hierarchical clustering 
```{r}
wisc.dist <- dist(scale(wisconsin.data))
```

```{r}
wisc.hclust <- hclust(wisc.dist)
plot(wisc.hclust)
```
```{r}
table(cutree(wisc.hclust, k=2))
```

## Combing different methods 
Take your pca data and then cluster this refined the clustering you;ll achieve. The idea here is that I can take my new variables (The PCs) that are better descriptors of the dataset than the origina features and find better clustering 

```{r}
pc.dist <- dist(wisc.pr$x[,1:3])
wisc.pr.hclust <- hclust(pc.dist, method = "ward.D2")
plot(wisc.pr.hclust)
```

```{r}
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)

table(grps, diagnosis)

```

```{r}
ggplot(wisc.pr$x) +
  aes(PC1, PC2) +
  geom_point(col=grps)
```

cluster 1 has 179 malignant while cluster. seems to be dominated by beningn 333

Sensitivity calucaltion 
```{r}
179/(179+34)
```

Specificity 
```{r}
333/(333 + 24)
```

```{r}
#url <- "new_samples.csv"
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```

```{r}
plot(wisc.pr$x[,1:2], col=grps)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```
